---
title: "Proposal title"
subtitle: "Proposal"
author: 
  - name: "Team Name"
    affiliations:
      - name: "School of Information, University of Arizona"
description: "Project description"
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: FALSE
  message: FALSE
---

```{r}
#| label: load-pkgs
#| message: false

if (!require("pacman")) 
  install.packages("pacman")
# use this line for installing/loading
pacman::p_load(readr,
               dplyr,
               ggplot2,
               scico,
               here,
               tidyverse,
               ggrepel,
               devtools,
               ggridges,
               dsbox,
               fs)
#library(tidyverse)
```

## Dataset

```{r}
#| label: load-dataset
#| message: false
#| warning: false
coffee_survey <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-05-14/coffee_survey.csv')


#glimpse(coffee_survey)
```

## Dataset Description

-   The coffee survey was collected from people participated in a YouTube event called the "Great American Coffee Taste Test". This event was primairly to taste test 4 different types of coffee, while the entire survey collected a wide variety of information. It has a total of 4042 participants/observations (rows), and 57 variables (columns).

-   I chose this data set because I am an avid coffee drinker and would like to know what other fellow coffee lovers like to drink, and if they spend as much money on coffee as I do. I also used to work at a local Mexican style coffee shop named Tierra Mia in the Los Angles area. Exploring this data set would just give me insight into the community that I would like to have.

## Questions

The two questions you want to answer.

Q1: Which age group spends the most money on coffee per month?

Q2: What is the favorite kinds of coffee for each age group?

## Analysis plan

-   I am planning on creating either a heat map like plot of the with count as the gradient or stacked bar charts
-   The variables I need to use are listed below along with the question that was asked to participants in the survey:
    -   total_spend: In total, how much money do you typically spend on coffee in a month?
    -   favorite: What is your favorite coffee drink?
    -   age: What is your age?
    -   Will have to create a count of total_spend variable per age group and count of favorite drinks per age group.

Task List:

| Task Name    | Status       | Assignee | Due   | Priority | Summary |
|--------------|--------------|----------|-------|----------|---------|
| Import Data  | compelte     | luis     | 05/29 | Low      |         |
| Explore Data | In Progress  | luis     | 06/01 | Moderate |         |
| Q1           | Not Started  | luis     | 06/03 | High     |         |
| Q2           | Not Started  | luis     | 06/05 | High     |         |
| Write Up     | Not Started  | luis     | 06/07 | High     |         |
| Presentation | Not Started  | luis     | 06/10 | High     |         |
| Final Review | Not Started  | luis     | 06/15 | Moderate |         |
