---
title: "Coffee Survey Proposal"
subtitle: "Proposal"
author: 
  - name: "Stats for Stacks"
    affiliations:
      - name: "School of Information, University of Arizona"
description: "Project description: How much money do different age groups spend on coffee? How do coffee type preferences differ by age group?"
format:
  html:
    code-tools: true
    code-overflow: wrap
    code-line-numbers: true
    embed-resources: true
editor: visual
code-annotations: hover
execute:
  warning: FALSE
  message: FALSE
---

```{r}
#| label: load-pkgs
#| warning: FALSE
#| message: FALSE
#| 
if (!require("pacman")) 
  install.packages("pacman")
# use this line for installing/loading
pacman::p_load(readr,dplyr,ggplot2,scico,here,tidyverse,ggrepel,devtools,
               ggridges,dsbox,fs,janitor)

```

## Dataset

```{r}
#| label: load-dataset
#| message: false
#| warning: false
#| 
coffee_survey <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-05-14/coffee_survey.csv')

glimpse(coffee_survey)

coffee_survey |> 
  colnames() |> 
  cat(sep = "\n")

coffee_survey <- coffee_survey|>
  janitor::clean_names() 
  
unique(coffee_survey$age)
ggplot(coffee_survey)+geom_bar(aes(age))+ 
  coord_flip(clip = "off")+labs(title='age')
unique(coffee_survey$favorite)
ggplot(coffee_survey)+geom_bar(aes(favorite))+ 
  coord_flip(clip = "off")+labs(title='favorite')
unique(coffee_survey$gender)
ggplot(coffee_survey)+geom_bar(aes(gender))+ 
  coord_flip(clip = "off")+labs(title='gender')
unique(coffee_survey$ethnicity_race)
ggplot(coffee_survey)+geom_bar(aes(ethnicity_race))+ 
  coord_flip(clip = "off")+labs(title='ethnicity_race')
unique(coffee_survey$education_level)
ggplot(coffee_survey)+geom_bar(aes(education_level))+ 
  coord_flip(clip = "off")+labs(title='education_level')
unique(coffee_survey$employment_status)
ggplot(coffee_survey)+geom_bar(aes(employment_status))+ 
  coord_flip(clip = "off")+labs(title='employment_status')
unique(coffee_survey$total_spend)
ggplot(coffee_survey)+geom_bar(aes(total_spend))+ 
  coord_flip(clip = "off")+labs(title='total_spend')
unique(coffee_survey$political_affiliation)
ggplot(coffee_survey)+geom_bar(aes(political_affiliation))+ 
  coord_flip(clip = "off")+labs(title='political_affiliation')
```

## Dataset Description

-   The coffee survey was collected from people participated in a YouTube event called the "Great American Coffee Taste Test". This event was primairly to taste test 4 different types of coffee, while the entire survey collected a wide variety of information. It has a total of 4042 participants/observations (rows), and 57 variables (columns).

-   I chose this data set because I am an avid coffee drinker and would like to know what other fellow coffee lovers like to drink, and if they spend as much money on coffee as I do. I also used to work at a local Mexican style coffee shop named Tierra Mia in the Los Angles area. Exploring this data set would just give me insight into the community that I would like to have.

## Questions

The two questions you want to answer.

Q1: Which age group spends the most money on coffee per month? There are 7 age groups. 5 range from 18-65. One is \<18, and another is \>65. Effectively covering ages from all walks of life.

Q2: What is the favorite kinds of coffee for each age group, gender, education level, employment status, and political affiliations, and ethnicity/race?

## Analysis plan

-   I am planning on creating either a heat map like plot of the with count as the gradient or stacked bar charts
-   The variables I need to use are listed below along with the question that was asked to participants in the survey:
    -   total_spend: In total, how much money do you typically spend on coffee in a month?
    -   favorite: What is your favorite coffee drink?
    -   age: What is your age?
    -   employment_status
    -   education_level
    -   gender
    -   political_affiliation
    -   ethnicity_race
    -   Will have to create a count of total_spend variable per age group and count of favorite drinks per age group.

Task List:

| Task Name    | Status      | Assignee | Due   | Priority | Summary                                                                               |
|--------------|-------------|----------|-------|----------|---------------------------------------------------------------------------------------|
| Import Data  | complete    | luis     | 05/29 | Low      | use read_csv to import data from the tidyTuesday github repo                          |
| Explore Data | complete    | luis     | 06/01 | Moderate | plot counts for each of the categorical variables to get a feel for whats in the data |
| Q1           | In Progress | luis     | 06/03 | High     |                                                                                       |
| Q2           | In Progress | luis     | 06/05 | High     |                                                                                       |
| Write Up     | Not Started | luis     | 06/07 | High     |                                                                                       |
| Presentation | Not Started | luis     | 06/10 | High     |                                                                                       |
| Final Review | Not Started | luis     | 06/15 | Moderate |                                                                                       |
